SMT-Boosted Security Types for Low-Level MPC

### 2025-05-10 esop25/new-cpj.tex
checkpoint

The primitive typing rules are precise and expressive and reflect no particular
implementation strategies, they just track dependencies as reflected in types.
The most important rules are for binary operators, where dependencies merge
in interesting ways especially with summation. In particular, if two
randomized values are summed, they remain randomized modulo disjointness
of randomness. Otherwise, overlapping randomness is merged into the dependencies.
For example:
\begin{eqnarray*}
& (\sx{w}{\cid_1} - \rx{w}{\cid_1}) + (\sx{w}{\cid_2} - \rx{w}{\cid_2}) : \\
& \dty{\setit{\rx{w}{\cid_1},\rx{w}{\cid_2}}}{\setit{\sx{w}{\cid_2},\sx{w}{\cid_1}}}
\end{eqnarray*}
but:
$$
(\sx{w}{\cid_1} - \rx{w}{\cid_1}) + \rx{w}{\cid_1} :
\dty{\varnothing}{\setit{\rx{w}{\cid_1},\sx{w}{\cid_1}}}
$$

\begin{definition}[Type Join]
  Given $\dty{R_1}{\eset_1}$ and $\dty{R_2}{\eset_2}$, define
  $\dty{R_1}{\eset_1} \tjoin \dty{R_2}{\eset_2} \defeq \dty{R}{\eset}$ where
  $R = (R_1 - (R_2 \cup \eset_2) \cup (R_2 - (R_1 \cup \eset_1))$ 
  and $\eset = (R_1 \cup R_2 \cup \eset_1 \cup \eset_2) - R $.
  %\begin{eqnarray*}
  %  R &\defeq& (R_1 - (R_2 \cup \eset_2) \cup (R_2 - (R_1 \cup \eset_1))\\
  %  \eset &\defeq& (R_1 \cup R_2 \cup \eset_1 \cup \eset_2) - R 
  %\end{eqnarray*}
\end{definition}

>>>>

The primitive typing rules are precise and expressive and reflect no particular
implementation strategies, they just track dependencies as reflected in types.
The most important rules are for binary operators, where dependencies merge
in interesting ways especially with summation. In particular, if two
randomized values are summed, they remain randomized modulo disjointness
of randomness. Otherwise, overlapping randomness is merged into the dependencies.
For example:
\begin{eqnarray*}
& (\sx{w}{\cid_1} - \rx{w}{\cid_1}) + (\sx{w}{\cid_2} - \rx{w}{\cid_2}) : \\
& \dty{\setit{\rx{w}{\cid_1},\rx{w}{\cid_2}}}{\setit{\sx{w}{\cid_2},\sx{w}{\cid_1}}}
\end{eqnarray*}
but:
$$
(\sx{w}{\cid_1} - \rx{w}{\cid_1}) + \rx{w}{\cid_1} :
\dty{\varnothing}{\setit{\rx{w}{\cid_1},\sx{w}{\cid_1}}}
$$
To join types appropriately, we make the following definition.
\begin{definition}[Type Join]
  Given $\dty{R_1}{\eset_1}$ and $\dty{R_2}{\eset_2}$, define
  $\dty{R_1}{\eset_1} \tjoin \dty{R_2}{\eset_2} \defeq \dty{R}{\eset}$ where
  $R = (R_1 - (R_2 \cup \eset_2) \cup (R_2 - (R_1 \cup \eset_1))$ 
  and $\eset = (R_1 \cup R_2 \cup \eset_1 \cup \eset_2) - R $.
  %\begin{eqnarray*}
  %  R &\defeq& (R_1 - (R_2 \cup \eset_2) \cup (R_2 - (R_1 \cup \eset_1))\\
  %  \eset &\defeq& (R_1 \cup R_2 \cup \eset_1 \cup \eset_2) - R 
  %\end{eqnarray*}
\end{definition}
Then the rules for typing are as follows. The most interesting are the rules
for summation that allow joining types and possibly retaining uniformity,
and the rule for binary operations (including multiplication) that
revert to the approximation of distributions of dependent expressions.
